Outbreak Threshold Detection

### 2023-03-09 notebooks/Julia/simulation-detection.qmd
Start to add time-varying beta (sine-wave scaled) Need to look into what's going on as sometimes the population increases exponentially, but the VariableRateJump should only affect the S -> E process, not the births.

```{julia}
using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
using Catalyst, Random, Statistics, DataFrames, CairoMakie, AlgebraOfGraphics
using BenchmarkTools, ProgressMeter

@@

```{julia}
tmax = 2.0
δt = 0.1/365
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);

@@

μ = 1
β = 1000
σ = 365/8
γ = 365/5
p = (β, σ, γ, μ)
```

```{julia}
expose_rate(u, p, t) = p[1] * u[1] * u[2] / (u[1] + u[2] + u[3] + u[4])  # β * S * I / N
function expose_affect!(integrator)
    integrator.u[1] -= 1         # S -> S - 1
    integrator.u[2] += 1         # E -> E + 1
    nothing
end
expose_jump = ConstantRateJump(expose_rate, expose_affect!)

@@

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), gillespie_jumps...)
sol = solve(jump_prob, SSAStepper())
sol_array = zeros(6, length(sol))
sol_array[1:4, :] = Array(sol)
sol_array[5, :] = sum(sol_array[1:4, :], dims = 1)
sol_array[6, :] = sol.t;
```

@@

```{julia}
colors = ["dodgerblue4", "firebrick3", "chocolate2", "green", "purple"]
sir_plot = data(stacked_df) *
    mapping(
        :time => "Time (Years)", :value => "Number",
        color = :variable => sorter("S", "E", "I", "R", "N") => "State"
    ) *
    visual(Lines, linewidth = 4)
draw(sir_plot; palettes = (; color = colors))
```

```{julia}
function run_sir_sim(; u₀, tspan, p)
    prob = DiscreteProblem(u₀, tspan, p)
    jump_prob = JumpProblem(prob, Direct(), gillespie_jumps..., save_positions = (false, false))
    jump_sol = solve(jump_prob, SSAStepper(), saveat = δt)
    return jump_sol;
end
```

@@

axislegend(ax)

>>>>

```{julia}
using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
using Catalyst, Random, Statistics, DataFrames, DataFramesMeta
using CairoMakie, AlgebraOfGraphics
using BenchmarkTools, ProgressMeter

@@

```{julia}
tmax = 10.0
δt = 0.1/365
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);

@@

μ = 1
β = 1000
σ = 365/8
γ = 365/5
ε = 100
p = (β, σ, γ, μ, ε)
```

```{julia}
x = 0:0.1:10
# 0.5 .* (sin.(2π .* x) .+ 1)
fig = lines(x, 0.5 .* (sin.(pi .* x ./ 2) .+ 1))
vlines!(collect(0:1:10); color = (:red, 0.5), linestyle = :dash)
fig
```

```{julia}
# 0.5 .* (sin.(2π .* x) .+ 1)
expose_rate(u, p, t) = p[1] * 0.5 * (sin(2π * t) + 1) * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])  # β * ϕ * S * I / N
function expose_affect!(integrator)
    integrator.u[1] -= 1         # S -> S - 1
    integrator.u[2] += 1         # E -> E + 1
    nothing
end

# We require that rate1(u,p,s) <= urate(u,p,s)
# for t <= s <= t + rateinterval(u,p,t)
rateinterval(u, p, t) = typemax(t)
urate(u, p, t) = p[1] * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])

# Optionally, we can give a lower bound over the same interval.
# This may boost computational performance.
lrate(u, p, t) = p[1] * u[1] * u[2]/ (u[1] + u[2] + u[3] + u[4])

# now we construct the bounded VariableRateJump
expose_jump = VariableRateJump(expose_rate, expose_affect!; lrate, urate, rateinterval)

# import_expose_rate(u, p, t) = p[5]
# function import_expose_affect!(integrator)
#     integrator.u[3] += 1         # I -> I + 1
#     nothing
# end
# import_expose_jump = ConstantRateJump(import_expose_rate, import_expose_affect!)

@@

```{julia}
# Constant jumps are ordered ahead of VariableJumps, but otherwise retain the ordering
dep_graph = [
    [1], [2], [3], [4], [5], [6], [7],
    [1, 2, 3, 4, 5, 6, 7, 8] # Exposure jump
    ]

prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Coevolve(), gillespie_jumps...; dep_graph)
sol = solve(jump_prob, SSAStepper())
sol_array = zeros(6, length(sol))
sol_array[1:4, :] = Array(sol)[1:4, :]
sol_array[5, :] = sum(sol_array[1:4, :], dims = 1)
sol_array[6, :] = sol.t
```

@@

```{julia}
colors = ["dodgerblue4", "firebrick3", "chocolate2", "green", "purple"]
sir_plot = data(stacked_df) *
    mapping(
        :time => "Time (Years)", :value => "Number",
        color = :variable => sorter("S", "E", "I", "R", "N") => "State"
    ) *
    visual(Lines, linewidth = 4)

draw(
    sir_plot;
    palettes = (; color = colors),
    axis = (; limits = ((0.0, 2.0), nothing))
    )
```

```{julia}
function run_sir_sim(; u₀, tspan, p)
    prob = DiscreteProblem(u₀, tspan, p)
    jump_prob = JumpProblem(
        prob, Coevolve(), gillespie_jumps..., save_positions = (false, false);
        dep_graph
        )
    jump_sol = solve(jump_prob, SSAStepper(), saveat = δt)
    return jump_sol;
end
```

@@

Legend(sir_sim_plot[1, 2], ax, "State")
